---
title: "Rworksheet_rabago#3a"
author: "James Bryan Rabago"
date: "2024-10-01"
output: pdf_document
---

```{r}
LETTERS
```
1.a. You need to produce a vector that contains the first 11 letters.
```{r}
f11 <- LETTERS [1:11]
f11
```

1.b. Produce a vector that contains the odd numbered letters.
```{r}
odd <- LETTERS [seq(1,26,by = 2)]
odd
```

1.c. Produce a vector that contains the vowels
```{r}
vowels <- LETTERS[c(1,5,9,15,21)]
vowels
```

1.d. Produce a vector that contains the last 5 lowercase letters.
```{r}
last <- letters [22:26]
last

```
1.e.Produce a vector that contains letters between 15 to 24 letters in lowercase.

```{r}
l15to24 <- letters [15:24]
l15to24
```



2.a.
```{r}
city <- c("Tuguegarao City", "Manila", "Iloilo City", "Tacloban", "Samal Island", "Davao City")
city
```

2.b.
```{r}
temp <- c(42, 39, 34, 34, 30, 27)
temp
```

2.c.
```{r}
citytemp <- data.frame(city,temp)
citytemp
```


2.d.
```{r}
names(citytemp) <- c("City","Temperature")
citytemp
```
```

2.e
```{r}
str(citytemp)
```
-the dataframe citytemp has 6 observations and 2 variables
the City column is a character vector(chr) while the Temperature column is a numeric(num).


2.f.
```{r}
citytemp[3:4, ]
```

2.g.

```{r}
citytemp[which.max(citytemp$Temperature),]
```


```{r}
citytemp[which.min(citytemp$Temperature),]
```

Tuguegarao City has the highest temperature city and Davao City has the lowest


2.a.
```{r}
mat<- matrix(c(1:8,11:14),ncol=4, nrow=3)
mat
```

2.b.
```{r}
mat2x <- mat * 2
mat2x
```

2.c.
```{r}
mat[2, ]
```

2.d.
```{r}
mat[1:2, 3:4]
```

2.e
```{r}
mat[3, 2:3]
```

2.f
```{r}
mat[ ,4]
```

2.g
```{r}
rownames(mat2x) <- c("isa","dalawa","tatlo")
colnames(mat2x) <- c("uno","dos","tres","quatro")
mat2x
```

2.h

```{r}
dim(mat) <- c(6,2)
mat
```


3.a.
```{r}
values <- rep(c(1,2,3,6,7,8,9,0,3,4,5,1), each=2)
myarray <- array(values, dim = c(2,4,3))
myarray
```

3.b.
```{r}
dim(myarray)
```

2 4 3 is the result


3.c
```{r}
dimnames(myarray) <- list(c("a","b"),c("A","B","C","D"),c("1st Dimensional Array","2nd Dimensional Array","3rd Dimensional Array"))
myarray
```

