---
title: "Rworksheet_Rabago#4c"
author: "James Bryan Rabago"
date: "2024-11-05"
output: pdf_document
---


```{r}
library(ggplot2)
write.csv(mpg, "mpg.csv", row.names = FALSE)
```


1.a.
```{r}
mpg <- read.csv("mpg.csv")
mpg
```

1.b.
-The categorical variables in the mpg dataset are manufacturer, model, trans, drv, fl, and class.



1.c.
-The continuous variables in the mpg dataset are displ, cty, and hwy.


2.a.
```{r}
library(dplyr)

manumodel<- mpg %>%
  group_by(manufacturer) %>%
  summarise(num_models = n_distinct(model)) %>%
  arrange(desc(num_models))

manumodel

```
2.b.
```{r}
barplot(manumodel$num_models,
        names.arg = manumodel$manufacturer,
        las = 2,col = "blue",
        main = "Number of Unique Models by Manufacturer",
        xlab = "Manufacturer",
        ylab = "Number of Unique Models")

```
```{r}
library(ggplot2)


ggplot(manumodel, aes(x = reorder(manufacturer, num_models), y = num_models)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  coord_flip() +  # Flip coordinates for better readability
  labs(title = "Number of Unique Models by Manufacturer",
       x = "Manufacturer",
       y = "Number of Unique Models") +
  theme_minimal()

```
2.a.
```{r}
ggplot(mpg, aes(model, manufacturer)) +
  geom_point()

```
The code creates a scatter plot showing car models and their manufacturers. However, this plot isn't very helpful because many models come from the same manufacturer, causing the points to overlap and making it hard to see individual models. Also, both the x-axis and y-axis use categories, which doesn't work well for showing relationships. To make the data clearer, using bar plots or boxplots would be more effective.


2.b.
The plot from `ggplot(mpg, aes(model, manufacturer)) + geom_point()` is not very useful due to overlapping points and categorical axes, making it hard to interpret. To improve it, you could create a bar plot showing how many models each manufacturer has, which highlights those with more variety. Alternatively, using boxplots to compare city or highway mileage by manufacturer would better illustrate fuel efficiency. These changes would make the visualizations more informative.



3.
```{r}

library(ggplot2)
library(dplyr)

top20 <- mpg %>%
  arrange(desc(cty)) %>%
  head(20)

ggplot(top20, aes(x = model, y = year)) +
  geom_point(color = "darkblue", size = 3) +
  coord_flip() +  
  labs(title = "Top 20 Models by Year of Manufacture",
       x = "Car Model",
       y = "Year") +
theme_minimal()

```

4
```{r}
library(dplyr)
library(ggplot2)


modelcount <- mpg %>%
  group_by(model) %>%
  summarise(num_cars = n()) %>%
  arrange(desc(num_cars))


modelcount
```
4.a.
```{r}
top20 <- modelcount %>%
  top_n(20, num_cars)

ggplot(top20, aes(x = reorder(model, num_cars), y = num_cars)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Top 20 Car Models by Number of Cars",
       x = "Car Model",
       y = "Number of Cars") +
  theme_minimal()

```
4.b.
```{r}
ggplot(top20, aes(x = reorder(model, num_cars), y = num_cars)) +
  geom_bar(stat = "identity", fill = "red") +
  coord_flip() +
  labs(title = "Top 20 Car Models by Number of Cars (Horizontal)",
       x = "Number of Cars",
       y = "Car Model") +
  theme_minimal()

```

5.
```{r}
library(ggplot2)

ggplot(mpg, aes(x = cyl, y = displ, color = displ)) +
  geom_point(size = 3) +
  labs(title = "Relationship between No. of Cylinders and Engine Displacement",
       x = "Number of Cylinders",
       y = "Engine Displacement (litres)") +
  theme_minimal() +
  scale_color_gradient(low = "blue", high = "red")  

```
6
```{r}
library(ggplot2)

ggplot(mpg, aes(x = displ, y = hwy, color = cty)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(title = "Relationship between Engine Displacement and Highway MPG",
       x = "Engine Displacement (litres)",
       y = "Highway MPG") +
  theme_minimal() +
  scale_color_gradient(low = "blue", high = "red")  # Color gradient from blue to red

```

-When you run the code, it generates a scatter plot showing the relationship between engine displacement and highway miles per gallon (MPG), with points colored by city MPG. Typically, you'll observe a negative trend: as engine displacement increases, highway MPG decreases, since larger engines tend to consume more fuel. The color coding adds context, with blue points (lower city MPG) often linked to higher displacement and lower highway MPG, while red points (higher city MPG) cluster at lower displacement and higher highway MPG. This illustrates the trade-off between engine size and fuel efficiency.




