---
title: "Rworksheet_Rabago#4c"
author: "James Bryan Rabago"
date: "2024-11-05"
output: pdf_document
---


```{r}
library(ggplot2)
write.csv(mpg, "mpg.csv", row.names = FALSE)
```


1.a.
```{r}
mpg <- read.csv("mpg.csv")
mpg
```

1.b.
-The categorical variables in the mpg dataset are manufacturer, model, trans, drv, fl, and class.



1.c.
-The continuous variables in the mpg dataset are displ, cty, and hwy.


2.a.
```{r}
library(dplyr)

manumodel<- mpg %>%
  group_by(manufacturer) %>%
  summarise(num_models = n_distinct(model)) %>%
  arrange(desc(num_models))

manumodel

```
2.b.
```{r}
barplot(manumodel$num_models,
        names.arg = manumodel$manufacturer,
        las = 2,col = "blue",
        main = "Number of Unique Models by Manufacturer",
        xlab = "Manufacturer",
        ylab = "Number of Unique Models")

```
```{r}
library(ggplot2)


ggplot(manumodel, aes(x = reorder(manufacturer, num_models), y = num_models)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  coord_flip() +  # Flip coordinates for better readability
  labs(title = "Number of Unique Models by Manufacturer",
       x = "Manufacturer",
       y = "Number of Unique Models") +
  theme_minimal()

```
2.a.
```{r}
ggplot(mpg, aes(model, manufacturer)) +
  geom_point()

```
The code creates a scatter plot showing car models and their manufacturers. However, this plot isn't very helpful because many models come from the same manufacturer, causing the points to overlap and making it hard to see individual models. Also, both the x-axis and y-axis use categories, which doesn't work well for showing relationships. To make the data clearer, using bar plots or boxplots would be more effective.


2.b.
The plot from `ggplot(mpg, aes(model, manufacturer)) + geom_point()` is not very useful due to overlapping points and categorical axes, making it hard to interpret. To improve it, you could create a bar plot showing how many models each manufacturer has, which highlights those with more variety. Alternatively, using boxplots to compare city or highway mileage by manufacturer would better illustrate fuel efficiency. These changes would make the visualizations more informative.



3.
```{r}

library(ggplot2)
library(dplyr)

top20 <- mpg %>%
  arrange(desc(cty)) %>%
  head(20)

ggplot(top20, aes(x = model, y = year)) +
  geom_point(color = "darkblue", size = 3) +
  coord_flip() +  
  labs(title = "Top 20 Models by Year of Manufacture",
       x = "Car Model",
       y = "Year") +
theme_minimal()

```

4
```{r}
library(dplyr)
library(ggplot2)


modelcount <- mpg %>%
  group_by(model) %>%
  summarise(num_cars = n()) %>%
  arrange(desc(num_cars))


modelcount
```
4.a.
```{r}
top20 <- modelcount %>%
  top_n(20, num_cars)

ggplot(top20, aes(x = reorder(model, num_cars), y = num_cars)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Top 20 Car Models by Number of Cars",
       x = "Car Model",
       y = "Number of Cars") +
  theme_minimal()

```
4.b.
```{r}
ggplot(top20, aes(x = reorder(model, num_cars), y = num_cars)) +
  geom_bar(stat = "identity", fill = "red") +
  coord_flip() +
  labs(title = "Top 20 Car Models by Number of Cars (Horizontal)",
       x = "Number of Cars",
       y = "Car Model") +
  theme_minimal()

```

5.
```{r}
library(ggplot2)

ggplot(mpg, aes(x = cyl, y = displ, color = displ)) +
  geom_point(size = 3) +
  labs(title = "Relationship between No. of Cylinders and Engine Displacement",
       x = "Number of Cylinders",
       y = "Engine Displacement (litres)") +
  theme_minimal() +
  scale_color_gradient(low = "blue", high = "red")  

```


6.1. The scatter plot shows a negative correlation between engine displacement (displ) and highway miles per gallon (hwy), suggesting that as engine displacement increases, highway MPG generally decreases. This relationship occurs because larger engines usually burn more fuel, resulting in reduced fuel efficiency on highways, highlighting the balance between engine size and fuel economy in vehicles.

```{r}
library(ggplot2)



ggplot(mpg, aes(x = displ, y = hwy, color = displ)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(title = "Relationship between Engine Displacement and Highway MPG",
       x = "Engine Displacement (Liters)",
       y = "Highway MPG",
       color = "Engine Displacement") +
  theme_minimal()

```
6.2. 
```{r}
traffic <- read.csv("/cloud/project/Worksheet4/traffic.csv")


```
a. 
```{r}

num_observations <- nrow(traffic)
print(paste("Number of observations:", num_observations))

variables <- colnames(traffic)
print("Variables in the dataset:")
print(variables)

```
b. 
```{r}


junction_subsets <- split(traffic, traffic$Junction)


```
c.
```{r}
library(ggplot2)
library(dplyr)


traffic$DateTime <- as.POSIXct(traffic$DateTime, format="%Y-%m-%d %H:%M:%S")

ggplot(traffic, aes(x = DateTime, y = Vehicles, color = factor(Junction))) +
  geom_line() +
  labs(title = "Traffic Volume Over Time by Junction",
       x = "Date and Time",
       y = "Number of Vehicles",
       color = "Junction") +
  theme_minimal() +
  facet_wrap(~ Junction, scales = "free_y")

```
7. 
```{r}
library(readxl)

alexa <- read_excel("/cloud/project/Worksheet4/alexa_file.xlsx")

```

a. 
```{r}
n <- nrow(alexa)
print(paste("Number of observations:", n))

v <- colnames(alexa)
print("Variables in the dataset:")
print(variables)
```
b. 
```{r}
library(dplyr)


variation_counts <- alexa %>%
  group_by(variation) %>%
  summarise(total = n())

print(variation_counts)

```
c.The plot shows which product variations are most popular, with some clearly preferred over others. It highlights what consumers like the most.
```{r}
ggplot(variation_counts, aes(x = reorder(variation, -total), y = total, fill = variation)) +
  geom_bar(stat = "identity") +
  labs(title = "Count of Each Product Variation",
       x = "Product Variation",
       y = "Total Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(fill = FALSE)
```
d. 
```{r}
alexa$date <- as.Date(alexa$date)

daily_reviews <- alexa %>%
  group_by(date) %>%
  summarise(total_verified_reviews = sum(feedback))

ggplot(daily_reviews, aes(x = date, y = total_verified_reviews)) +
  geom_line(color = "blue") +
  labs(title = "Number of Verified Reviews Over Time",
       x = "Date",
       y = "Total Verified Reviews") +
  theme_minimal()
```
e. 
```{r}
library(dplyr)
library(ggplot2)

variation_ratings <- alexa %>%
  group_by(variation) %>%
  summarise(average_rating = mean(rating, na.rm = TRUE)) %>%

arrange(desc(average_rating))

variation_ratings
```


```{r}
highvar <- variation_ratings %>%
  slice(1)
highvar
```


```{r}
ggplot(variation_ratings, aes(x = reorder(variation, -average_rating), y = average_rating, fill = variation)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Rating by Product Variation",
       x = "Product Variation",
       y = "Average Rating") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(fill = FALSE)
```

