---
title: "Rworksheet_Rabago#4a"
author: "James Bryan Rabago"
date: "2024-10-15"
output: pdf_document
---
1.
```{r}
data <- data.frame(
ShoeSize = c(6.5,9.0,8.5,8.5,10.5,7.0,9.5,9.0,13.0,7.5,10.5,8.5,12.0,10.5,13.0,11.5,8.5,5.0,10.0,6.5,7.5,8.5,10.5,8.5,10.5,11.0,9.0,13.0),
Height = c(66.0,68.0,64.5,65.0,70.0,64.0,70.0,71.0,72.0,64.0,74.5,67.0,71.0,71.0,77.0,72.0,59.0,62.0,72.0,66.0,64.0,67.0,73.0,69.0,72.0,70.0,69.0,70.0),
Gender = c("F","F","F","F","M","F","F","F","M","F","M","F","M","M","M","M",
           "F","F","M","F","F","M","M","F","M","M","M","M") )

data
```

1.a.


The data contains three variables. The first variable is shoe size, which is numeric and shows the respondents' shoe sizes. The second variable is height, also numeric, representing the respondents' height in inches. The third variable is gender, which is categorical and represented as "M" for male and "F" for female.


1.b.
```{r}
fshoesize <- subset(data,Gender == "F",select = c(ShoeSize,Height))
mshoesize <- subset(data,Gender == "M",select = c(ShoeSize,Height))

fshoesize
mshoesize
```


1.c.
```{r}
meanshoesize <- mean(data$ShoeSize)
meanheight <- mean(data$Height)

meanshoesize
meanheight
```
1.d.


-There is no relationship between shoe size and height because they are separate physical traits. Even though both vary from person to person, one does not directly affect the other. People of the same height can have different shoe sizes, and vice versa.




2.
```{r}
months_vector <- c("March", "April", "January", "November", "January",
                   "September", "October", "September", "November", "August",
                   "January", "November", "November", "February", "May", "August", 
                   "July", "December", "August", "August", "September", "November", 
                   "February", "April")


factor_months_vector <- factor(months_vector)


factor_months_vector
```
3.
```{r}
summary(months_vector)
summary(factor_months_vector)
```
4.
```{r}
direction <- c("East","West","North")
frequency <- c(1,4,3)

factor_data <- direction
new_order_data <- factor(factor_data, levels = c("East","West","North"))
new_order_data
```

5.a.
```{r}
import_march <- read.table("import_march.csv", header = TRUE, sep = ",")
```


5.b.
```{r}
import_march
```

6.a

```{r}
number <- readline(prompt= "Select number 1 to 50: ")
number

if(number < 1 && number > 50){
  print("The number selected is beyond the range of 1 to 50")
}else if (number == 20){print(paste("TRUE"))
}else {print(paste("The selected number is: ",number))}

```
7.a.

```{r}
price <- as.numeric(readline(prompt="Enter the price: "))

minimumbills <- function(price) {
billsused <- 0
if (!is.na(price) && price >= 1000) {
billsused <- billsused + price %/% 1000
price <- price %% 1000
}

if (!is.na(price) && price >= 500) {
billsused <- billsused + price %/% 500
price <- price %% 500
}
if (!is.na(price) && price >= 200) {
billsused <- billsused + price %/% 200
price <- price %% 200
}
if (!is.na(price) && price >= 100) {
billsused <- billsused + price %/% 100
price <- price %% 100
}
if (!is.na(price) && price >= 50) {
billsused <- billsused + price %/% 50
price <- price %% 50
}
return(billsused)
}
minimumbills(price)

```
8.a.
```{r}

students <- data.frame(
  Name = c("Annie", "Thea", "Steve", "Hanna"),
  Grade1 = c(85, 65, 75, 95),
  Grade2 = c(65, 75, 55, 75),
  Grade3 = c(85, 90, 80, 100),
  Grade4 = c(100, 90, 85, 90)
)


students

```
8.b.
```{r}

avgscores <- rowSums(students[,-1]) / (ncol(students) - 1)


if (any(avgscores > 90)) {
 
  highavg_names <- students$Name[avgscores > 90]
  highavg_scores <- avgscores[avgscores > 90]
  
  
  for (i in 1:length(highavg_names)) {
    cat(paste(highavg_names[i], "'s average grade this semester is ", highavg_scores[i], ".\n", sep = ""))
  }
} else {
  cat("No student's average grade is over 90.\n")
}
```

8.c.
```{r}

testaverages <- colSums(students[,-1]) / nrow(students)


if (any(testaverages < 80)) {
  difficult_tests <- which(testaverages < 80)
  cat(paste ("The", difficult_tests, "th test was difficult.\n"))
} else {
  cat("All tests had an average score of 80 or above.\n")
}

```
8.d.
```{r}

for (i in 1:nrow(students)) {
  highest_score <- students$Grade1[i]
  if (students$Grade2[i] > highest_score) highest_score <- students$Grade2[i]
  if (students$Grade3[i] > highest_score) highest_score <- students$Grade3[i]
  if (students$Grade4[i] > highest_score) highest_score <- students$Grade4[i]
  
  if (highest_score > 90) {
    cat(students$Name[i], "highest grade this semester is", highest_score, "\n")
  }
}

```

