---
title: "Rworksheet_rabago#2"
author: "James Bryan Rabago"
date: "2024-09-30"
output: html_document
---
1.
a.
```{r}
vector_a <- -5:5
vector_a
```
the ":" operator creates a sequence of numbers from the -5 to 5


b.
```{r}
x <- 1:7
x
```
the value of x contains a sequence of integers starting from 1 to 7


2.
```{r}
vectorseq <- seq(1,3, by = 0.2)
vectorseq
```
the output shows numbers ranging from 1 to 3 with a step sixe of 0.2


3.
```{r}
workages <- c(34,28,22,36,27,18,52,39,42,29,35,31,27,22,37,34,19,20,57,49,50,37,46,25,17,37,43,53,41,51,35,24,33,41,53,40,18,44,38,41,48,27,39,19,30,61,54,58,26,18)
workages
```
3.a.
```{r}
e3rd <- workages[3]
e3rd
```
3.b.
```{r}
e24 <- workages[c(2,4)]
e24
```
3.c.
```{r}
not412 <- workages[-c(4,12)]
not412
```
4.
```{r}
names <- c("first"=3,"second"=0,"third"=9)
```
4.a.the output is 3 and 9 which is the values of "first" and "third" elements
4.b.
```{r}
names
names[c("first","third")]
```
5.
```{r}
x <- -3:2
x
```
5.a.
after modifying the 2nd element which is -2 updated to 0


5.b.
```{r}
x[2] <- 0
x
```

6.
```{r}
month <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun")
price <- c(52.50, 57.25, 60.00, 65.00, 74.25, 54.00)
purchase <- c(25, 30, 40, 50, 10, 45)
```
6.a.
```{r}
fueldata <- data.frame(Month = month, Price = price, Purchase = purchase)
fueldata
```
6.b.
```{r}
avgfuel <- weighted.mean(price, purchase)
avgfuel
```
7.a.
```{r}
narivers <- c(length(rivers), sum(rivers), mean(rivers), median(rivers), var(rivers), sd(rivers), min(rivers), max(rivers))
narivers
```
7.b.
the output prints the length,sum,mean,median,variance,standard deviation,minimum and maximum of rivers

7.c
```{r}
narivers <- c(length(rivers), sum(rivers), mean(rivers), median(rivers), var(rivers), sd(rivers), min(rivers), max(rivers))
narivers
```

8.
```{r}
celdat <- data.frame(
  Power_Ranking <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25),
  Celebrity_Name <- c("Tom Cruise", "Rolling Stones", "Oprah Winfrey", "U2", "Tiger Woods", "Steven Spielberg",  "Howard Stern", "50 Cent", "Cast of the Sopranos", "Dan Brown", "Bruce Springsteen",  "Donald Trump", "Muhammad Ali", "Paul McCartney", "George Lucas", "Elton John", 
"David Letterman", "Phil Mickelson", "J.K Rowling", "Bradd Pitt", "Peter Jackson","Dr. Phil McGraw", "Jay Lenon", "Celine Dion", "Kobe Bryant"),
  Pay <- c(67, 90, 225, 110, 90, 332, 302, 41, 52, 88, 55, 44, 55, 40, 233, 34, 40, 47, 75, 39, 45, 32, 30, 33, 31)
)
celdat

```

8.b.
```{r}
Rowling <- which(Celebrity_Name == "J.K. Rowling")
Power_Ranking[Rowling] <- 15
Pay[Rowling] <- 90

Power_Ranking
Pay
```
8.c
```{r}
write.csv("celdat","celebrities_data.csv", row.names=FALSE)
```

```
8.d.
```{r}
c1020rows <- celdat[10:20,]
c1020rows
```


```{r}
save(c1020rows, file = "Ranks.RData")
c1020rows
```

e.The rows accessed include the 10th to the 20th celebrities from the original Forbes ranking table. After filtering, the subset was saved in an RData file format.


9.a.
```{r}
library(openxlsx)
hotels_data <- read.xlsx("hotels-vienna.xlsx")
hotels_data
```

9.b.

```{r}
dim(hotels_data)
```

9.c.
```{r}
selected <- hotels_data[, c("country", "neighbourhood", "price", "stars", "accommodation_type", "rating")]

selected
```

9.d.
```{r}
save(selected, file ="new.RData")
load("new.RData")
head(selected)
tail(selected)
```

10.a.
```{r}
vegetables <- list("Carrot","Potato","Onions","Broccoli","Lettuce","Capsicum","Cabbage","Eggplant","Squash","Celery")

vegetables
```

10.b.
```{r}
vegetables <- append(vegetables, list("Zuccini","Bell Pepper"))
vegetables
```

10.c.
```{r}
vegetables <- append(vegetables, list("Mushroom","Cauliflower","Kale","Radish"),after = 5)
vegetables

length(vegetables)
```
16 datapoints does the vegetables list have


10.d.
```{r}
vegetables <- vegetables[-c(5,10,15)]
vegetables

length(vegetables)
```

13 vegetables were left